Regression Analysis of Chicago Crime Volume by Ward using Public School Data

### 2018-11-08 Paper.tex
just backing up tex paper

\paragraph{Independent Variables}
Independent variables were then identified from the dataset, based on theory, prior work on this subject, and availability.  OLS necessitates that all variables used with the estimator must be stated numerically; therefore, all chosen variables are either numeric or able to be translated as such.\footnote{Dress Code is a dummy, binary variable.  Schools with a dress code are valued at 1, while those without a dress code are denoted by 0.  Therefore, each ward will have a dress code value of 0-1, depending on the mix of high schools in that ward.}\footnote{Survey variables result from surveys given to teachers and students to rank their school in five criteria:  supportive environment, involved families, school safety, collaborative teaching, and ambitious instruction.  Scoring is from 0-5 with 5 being the best.}  This precluded many free-format data fields consisting of comments or non-ordinal evaluation.  The chosen variables, and their descriptive statistics, can be seen in Table 2.  These consist of the initial variables that were evaluated for regression analysis, not the final model's variables.

@@

These candidate variables were then evaluated for multicollinearity, as correlated independent variables violate one of the basic assumptions of OLS.  This was completed visually (since the number of variables chosen was relatively small) utilizing R, a statistical software.  The resulting correlations can be seen in Figure 1.  This information was referenced as final variables were selected.

@@

\subsection{Variable Selection \& Model Estimation}

The model chosen to explain crime as a function of public school data was a regression using Ordinary Least Squares as the estimator.  The variables shown in Table 2 represent all of the possible attributes from the data that could theoretically serve as a predictor and that could be coded in a numeric format, as explained earlier.  In order to narrow these variables into an initial model, stepwise regression was utilized.  This was conducted using R, with a separate analysis conducted for each of the dependent variables:  total crime and total homicides.  Once completed, these selected variables were tested for multicollinearity and heteroscedasticity, using Variance Inflation Factors and White's test respectively, analyzed as a regression in SAS, and then subsequently worked into the best possible model given the data.

@@

The aforementioned variables were then run using \textit{PROC REG}, the regression command in SAS.  Upon evaluation, average ACT score and the leadership survey variables stood out with unacceptable high p-values of 0.90 and 0.19, respectively.  These variables were then dropped from the estimate, and the regression was recalculated.  The results can be seen in Table 3.

			& \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.} & \textbf{Minimum} & \\
			\hline

		\end{tabular}
		\caption{Coefficients, standard error, and p-values for linear regression of total crime}
	\end{table}
\end{center}

As can be seen above, all selected variables have acceptable p-values that fall below the commonly-held threshold of 0.05.  Furthermore, all variables included pass the White test for heteroscedasticity adjustment and have suitably low Variance Inflation Factors, as can be seen in Table 4.  For the White adjustment for heteroscedasticity shows that all of the included variables maintain an acceptably low p-value and standard error.  A Variance Inflation Factor of less than 5 is typically accepted as not indicating severe multicollinearity, and all of the included variables fall below this limit.

@@

OVERALL THE REGRESSION QUALITY

			& \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.} & \textbf{Minimum} & \\
			\hline

@@

BRIEF INTRO HERE

	\paragraph{Total Crime Regression}

	\paragraph{Total Homicide Regression}

>>>>

\paragraph{Independent Variables}
Independent variables were then identified from the dataset, based on theory, prior work on this subject, and availability.  OLS necessitates that all variables used with the estimator must be stated numerically; therefore, all chosen variables are either numeric or able to be translated as such.\footnote{Dress Code is a dummy, binary variable.  Schools with a dress code are valued at 1, while those without a dress code are denoted by 0.  Therefore, each ward will have a dress code value of 0-1, depending on the mix of high schools in that ward.}\footnote{Survey variables result from surveys given to teachers and students to rank their school in five criteria:  supportive environment, involved families, school safety, collaborative teaching, and ambitious instruction.  Scoring is from 0-5 with 5 being the best.}  This precluded many free-format data fields consisting of comments or non-ordinal evaluation.  The chosen variables, and their descriptive statistics, can be seen in Table 2.  These consist of the initial variables that were evaluated for regression analysis, not the final model's variables.
\par
These candidate variables were then evaluated for multicollinearity, as correlated independent variables violate one of the basic assumptions of OLS.  This was completed visually (since the number of variables chosen was relatively small) utilizing R, a statistical software.  The resulting correlations can be seen in Figure 1.  This information was referenced as final variables were selected.

@@

@@

\subsection{Model Specification, Variable Selection, \& Model Estimation}

The model chosen to explain crime as a function of public school data was a linear regression using Ordinary Least Squares as the estimator.  Exploratory analysis of the target variables and the literature review suggested that a linear relationship is the most fitting for this analysis.  Furthermore, each ward has essentially the same population, so total crime count by ward represents a consistent rate, eliminating the need to standardize it.

The variables shown in Table 2 represent all of the possible attributes from the data that could theoretically serve as a predictor and that could be coded in a numeric format, as explained earlier.  In order to narrow these variables into an initial model, stepwise regression was utilized.  This was conducted using R, with a separate analysis conducted for each of the dependent variables:  total crime and total homicides.  Once completed, these selected variables were tested for multicollinearity and heteroscedasticity, using Variance Inflation Factors and White's test respectively, analyzed as a regression in SAS, and then subsequently worked into the best possible model given the data.

@@

The aforementioned variables were then run using \textit{PROC REG}, the regression command in SAS.  Upon evaluation, average ACT score and the leadership survey variables stood out with unacceptably high p-values of 0.90 and 0.19, respectively.  These variables were then dropped from the estimate, and the regression was recalculated.  The resulting variable metrics are shown in Table 3, and the resulting estimate equation can be written as:

\begin{math}
TotalCrime = 2185 - 62.22*GraduationRate + 73.35*CollegeEnrollmentRate + 2451.65*DressCode + 138.44*DropoutRate
\end{math}

\begin{center}
	\begin{table}[h]
		\begin{tabular}{ c | c | c | c | c }

			& \textbf{Grad. Rate} & \textbf{College Enroll} & \textbf{Dress Code} & \textbf{Dropout Rate} \\
			\hline
			\textbf{Coefficient Estimate} & -62.22 & 73.35 & 2451.65 & 138.44 \\
			\textbf{Standard Error} & 31.16 & 29.95 & 844.39 & 47.29 \\
			\textbf{T-Statistic} & -2.00 & 2.45 & 2.90 & 2.93\\
			\textbf{P-Value} & 0.0522 & 0.0185 & 0.0058 & 0.0054		
		\end{tabular}
		\caption{Coefficients, standard error, and p-values for linear regression of total crime}
	\end{table}
\end{center}

The same variables shown in Table 3 were fit to total homicides as the target variable.  With less occurrences, and in some wards no occurrences, this model fitting was conducted as exploratory.  In order to build a more accurate regression, more robust homicide data is necessary.  The variable statistics can be seen in Table 4, and the resulting regression estimate was determined to be:

\begin{math}
TotalHomicides = 16.3 - 0.236*GraduationRate + 0.045*CollegeEnrollmentRate + 12.125*DressCode + 0.305*DropoutRate $$
\end{math}

\begin{center}
	\begin{table}[h]
		\begin{tabular}{ c | c | c | c | c }

			& \textbf{Grad. Rate} & \textbf{College Enroll} & \textbf{Dress Code} & \textbf{Dropout Rate} \\
			\hline
			\textbf{Coefficient Estimate} & -0.236 & 0.045 & 12.125 & 0.305 \\
			\textbf{Standard Error} & 0.161 & 0.155 & 4.375 & 0.245 \\
			\textbf{T-Statistic} & -1.46 & 0.29 & 2.77 & 1.25\\
			\textbf{P-Value} & 0.1505 & 0.7737 & 0.0082 & 0.2193		
		\end{tabular}
		\caption{Coefficients, standard error, and p-values for linear regression of total homicides}
	\end{table}
\end{center}

	\subsection{Total Crime Regression}

	\paragraph{Theory}
	When the equation estimate is viewed through the lens of theoretical relationship, it perhaps seems somewhat counterintuitive at first.  However, most of the variables' relationship to the number of total crimes can be rationalized.  It is important to note, however, that the presence of a relationship does not necessarily indication causation.  Schools with a lower graduation rates support the theory that students with less options for a future (like not possessing a high school diploma) may be more inclined to get involved with criminal activity.  This is reflected in that variable's negative coefficient.  Dress code should theoretically reduce crime by making gang affiliations less apparent, yet it has a positive relationship with crime count.  It could be the case that institution of a dress code is reactionary to crime; that is, schools in higher crime areas require a dress code in order to help with the issue.  Dropout rate has a positive relationship with crime, which furthers the argument that students with less options are perhaps more inclined to engage in crime.

@@

	College enrollment is the one variable that is perhaps not explainable on a theoretical level.  Common sense indicates that college-bound students are perhaps less engaged in criminal activity than non-college-bound students.  Similarly, higher-crime areas would be expected to have less families with a college graduate than lower-crime areas.  However, the variable is shown to be statistically significant, and further research may be necessary to explore this relationship.

	\paragraph{Fit}

	The linear model for total crime as a function of the aforementioned four variable exhibits a fairly reasonable level of fit.  It has a $ R^{2} $ value of 0.3154 and a $ \bar{R}^{2} $ of 0.2518.  Both metrics measure overall fit on a scale from 0-1, with $ \bar{R}^{2} $ ``penalizing" for the addition of more variables.  Since this is a multivariate regression, both metrics must be taken into account.  The values for these suggest that this is a model that is indeed indicative of a relationship.  The chosen variables maximize these metrics given the data at hand, since the variables were chosen utilizing a stepwise regression algorithm set to maximize $ \bar{R}^{2} $.
	\par

	A regression's F-value ajfl;fafjasflasjflajlf

	Furthermore, all variables in this regression estimate have p-values that are low enough to be considered statistically significant.  For this measure, a standard p-value of less than 0.05 is considered to be significant.  Since all of the variables exhibit this, the null hypothesis, which states that the coefficient is expected to be zero, is rejected.

	\paragraph{Potential Issues}All variables included in the total crime regression pass the White test for heteroscedasticity adjustment and have suitably low Variance Inflation Factors, as can be seen in Table 4.  For the White adjustment for heteroscedasticity shows that all of the included variables maintain an acceptably low p-value and standard error.  A Variance Inflation Factor of less than 5 is typically accepted as not indicating severe multicollinearity, and all of the included variables fall below this limit.

	\begin{center}
		\begin{table}[h]
			\begin{tabular}{ c | c | c | c | c }

				& \textbf{Coefficient Estimate} & \textbf{Adj. Std. Error} & \textbf{Adj. P-Value} & \textbf{VIF} \\
				\hline
				\textbf{Graduation Rate} & -62.22 & 27.57 & 0.029 & 2.12  \\
				\textbf{College Enrollment} & 73.35 & 28.55 & 0.014 & 3.38 \\
				\textbf{Dress Code} & 2451.65 & 718 & 0.001 & 1.08 \\
				\textbf{Dropout Rate} & 138.43 & 51.07 & 0.010 & 1.96

@@

	\subsection{Total Homicide Regression}

	\paragraph{Theory}
	When the equation estimate is viewed through the lens of theoretical relationship, it perhaps seems somewhat counterintuitive at first.  However, most of the variables' relationship to the number of total crimes can be rationalized.  It is important to note, however, that the presence of a relationship does not necessarily indication causation.  Schools with a lower graduation rates support the theory that students with less options for a future (like not possessing a high school diploma) may be more inclined to get involved with criminal activity.  This is reflected in that variable's negative coefficient.  Dress code should theoretically reduce crime by making gang affiliations less apparent, yet it has a positive relationship with crime count.  It could be the case that institution of a dress code is reactionary to crime; that is, schools in higher crime areas require a dress code in order to help with the issue.  Dropout rate has a positive relationship with crime, which furthers the argument that students with less options are perhaps more inclined to engage in crime.

	College enrollment is the one variable that is perhaps not explainable on a theoretical level.  Common sense indicates that college-bound students are perhaps less engaged in criminal activity than non-college-bound students.  Similarly, higher-crime areas would be expected to have less families with a college graduate than lower-crime areas.  However, the variable is shown to be statistically significant, and further research may be necessary to explore this relationship.

	\paragraph{Fit}

	r square and stuff

	A regression's F-value ajfl;fafjasflasjflajlf

	p-value

	\paragraph{Potential Issues}

	\begin{center}
	\begin{table}[h]
		\begin{tabular}{ c | c | c | c | c }

			& \textbf{Coefficient Estimate} & \textbf{Adj. Std. Error} & \textbf{Adj. P-Value} & \textbf{VIF} \\
			\hline
			\textbf{Graduation Rate} & -0.236 & 0.146 & 0.1122 & 2.12  \\
			\textbf{College Enrollment} & 0.045 & 0.132 & 0.7346 & 3.38 \\
			\textbf{Dress Code} & 12.125 & 3.153 & 0.0004 & 1.08 \\
			\textbf{Dropout Rate} & 0.305 & 0.201 & 0.1366 & 1.96

		\end{tabular}
		\caption{Heteroscedasticity and multicollinearity test results for linear regression model of total homicides}
	\end{table}
\end{center}
